\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage[greek,english]{babel}
  %option greek for \<euro>
  %option english (default language) for \<guillemotleft>, \<guillemotright>

\usepackage[latin1]{inputenc}
  %for \<onesuperior>, \<onequarter>, \<twosuperior>, \<onehalf>,
  %\<threesuperior>, \<threequarters>, \<degree>

\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<cent>, \<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{Lattice Properties}

\author{Viorel Preoteasa}

\maketitle

\begin{abstract}
This formalization introduces some algebraic structures
based on lattices and complete lattices. These results 
are used in some other formalizations and it seems natural
to collect them all in one formalization instead of
repeating them. The structures introduced here are
complete distributive lattices, modular and distributive
lattices, as well as lattice ordered groups. In addition 
to the results proved for the new lattices, this
formalization also introduces some theorems
about latices and distributive lattices in general.
\end{abstract}

\tableofcontents

\parindent 0pt\parskip 0.5ex

\section{Overview}

Section 2 introduces complete distributive lattices, complete
Boolean algebras, and proves the De Morgan's laws for arbitrary
conjunctions and disjunctions. Section 3 proves some properties
of complete lattices. The most important property is that 
a monotonic application which maps monotonic functions to
monotonic functions has the least fixpoint monotonic.
Section 4 introduces the classes inf, sup, and it redefines
the lattice class as subclass of inf and sup. This construction
is done in order to enable easier reuse of the inf and sup
operators. Some simplification lemmas are also proved in 
Section 4.
Section 5 introduces modular lattices and proves some properties
about them and about distributive lattices. The main result
of this section is that a lattice is distributive if and only
if it satisfies
$$\forall x \; y \; z: x \sqcap z = y \sqcap z \land  x \sqcup z = y \sqcup z \longrightarrow x = y$$
Section 6 introduces lattice ordered groups and some of their properties.
The most important is that they are distributive lattices, and
this property is proved using the results from Section 5.

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
