\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage[a4paper,margin=2.4cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{microtype}
\usepackage{url}

\urlstyle{rm}
\isabellestyle{it}
\setlength{\emergencystretch}{3em}

\begin{document}

\title{Vosper's Theorem and the Cauchy--Davenport Theorem via the Polynomial Method}
\author{Arthur F. Ramos \and David Barros Hulak \and Ruy J. G. B. de Queiroz}
\maketitle

\begin{abstract}
This entry formalizes the Cauchy--Davenport lower bound and Vosper's theorem
for sumsets over finite fields of prime cardinality. It combines a
polynomial-method proof of Cauchy--Davenport with a Davenport-transform proof of
Vosper's classification of the nontrivial equality case, and it provides
modular-representative lemmas connecting the abstract prime-field results to
normalized representatives of $\mathbb{Z}/p\mathbb{Z}$. AI assistance was used
for proof engineering. The final definitions, statements, and proofs are checked by
Isabelle.
\end{abstract}

\section{Overview}

For nonempty finite subsets $A$ and $B$ of a prime field $\mathbb{F}_p$, the
Cauchy--Davenport theorem gives the lower bound
\[
  |A + B| \geq \min(p, |A| + |B| - 1).
\]
Vosper's theorem refines the extremal case: if $|A|, |B| \geq 2$, the sumset is
not all of $\mathbb{F}_p$, and equality holds in the Cauchy--Davenport bound,
then both sets are arithmetic progressions with a common difference. These
results are standard cornerstones of additive combinatorics.

\section{Formalization}

The Isabelle/HOL development first establishes generic finite-sumset facts for
translations, complements, and cardinalities. The proof of Cauchy--Davenport is
then carried out by the polynomial method, using ordinary Isabelle polynomials
to package the bivariate degree argument over finite fields of prime
cardinality. Vosper's theorem is proved via Davenport transforms: the
formalization derives one-sided progression results under normalization
hypotheses and combines them to obtain the final shared-difference
classification. The entry also supplies concrete modular-translation facts over
normalized representatives of $\mathbb{Z}/p\mathbb{Z}$.

\section{Relation to existing AFP entries}

The Cauchy--Davenport bound has been formalized in the AFP before. The entry
\emph{Kneser's Theorem and the Cauchy--Davenport
Theorem}~\cite{Baksys_Kneser_CD_AFP} derives it as a corollary of Kneser's
addition theorem for abelian groups, and \emph{A Generalization of the
Cauchy--Davenport Theorem}~\cite{Baksys_Generalized_CD_AFP} formalizes DeVos's
generalization to (possibly non-abelian) groups. Both developments are set in an
abstract group structure and build on further AFP material.

The present entry is intended to complement, not duplicate, this work. Its
principal new contribution is \emph{Vosper's theorem}, the critical-pair
(inverse) theorem characterizing the equality case of the Cauchy--Davenport
bound, which neither of the entries above provides. The Cauchy--Davenport bound
itself is established here by a different route, the \emph{polynomial method}
(Combinatorial Nullstellensatz) following Alon, Nathanson, and
Ruzsa~\cite{AlonNathansonRuzsa1996}: a single explicit polynomial is shown to be
non-vanishing on a grid built from $A$ and $B$. To the best of our knowledge
this proof of Cauchy--Davenport is not otherwise present in the AFP. It is
carried out directly for finite fields of prime cardinality, modelled by the
\texttt{finite\_field} type class together with a primality hypothesis on the
field's cardinality, and it depends only on the Isabelle/HOL distribution
(\texttt{HOL-Computational\_Algebra} and \texttt{HOL-Number\_Theory}). Keeping
this self-contained prime-field bound as the base case lets the Vosper
development proceed without translating between the type-class and locale-based
group representations.

\section{Sources}

The Cauchy--Davenport proof follows the polynomial-method perspective of Alon,
Nathanson, and Ruzsa~\cite{AlonNathansonRuzsa1996}. The additive-combinatorics
background follows standard sources such as Nathanson~\cite{Nathanson1996} and
Tao and Vu~\cite{TaoVu2006}. The inverse theorem formalized here is Vosper's
critical-pair theorem for groups of prime order~\cite{Vosper1956}.

\tableofcontents

\sloppy
\input{session}
\fussy

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

