\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

\usepackage{amsfonts, amsmath, amssymb}
\usepackage{textcomp}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-style
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{The Banach--Tarski Paradox in Isabelle/HOL}
\author{Arthur F. Ramos \and David Barros Hulak \and Ruy J. G. B. de Queiroz}
\date{}
\maketitle

\begin{abstract}
We formalise in Isabelle/HOL the classical Banach--Tarski paradox for the
closed unit ball in three-dimensional Euclidean space.  The development
proves the free-group paradox, the free subgroup of \(SO(3)\),
the Hausdorff paradox away from a countable bad set, the absorption
arguments for the sphere and the origin, and the final finite-partition
theorem for the ball.

The proof follows the standard route: the paradoxical decomposition of the
free group on two generators \(F_2\), the existence of a free subgroup
\(F_2 \le SO(3)\) via the AFP \texttt{Free-Groups} entry, the Hausdorff
paradox for \(S^2 \setminus D\) where \(D\) is a countable set of fixed
points, and absorption arguments extending the partition to \(S^2\), then
radially to \(B^3 \setminus \{0\}\), and finally to \(B^3\) itself.
The argument crucially relies on the axiom of choice.
\end{abstract}

\section*{Background}

The Banach--Tarski paradox was proved by Banach and Tarski
\cite{BanachTarski1924}, building on Hausdorff's earlier paradoxical
decomposition of the sphere \cite{Hausdorff1914}.  The route formalised here
follows the standard free-group presentation from Wagon's expositions
\cite{Wagon1993,TomkowiczWagon2016} and can be compared with de Rauglaudre's
Coq formalisation \cite{deRauglaudre2017BanachTarski}.  The Isabelle
development uses Breitner's AFP entry on free groups
\cite{AFPFreeGroups}.

\section*{Formalisation Profile}

The generated Isabelle session is \texttt{Banach\_Tarski}.  It is based on the
HOL Analysis library, imports the AFP entry \texttt{Free-Groups}, and uses
the \texttt{Primes} theory for the arithmetic freeness
argument.  The development introduces no local axiomatization or oracle.
Its non-constructive steps use the standard Hilbert-choice operators of
HOL, made explicit in the orbit-representative construction and inverse-map
uses.

The checked target is Isabelle 2025-2 with AFP 2026-04-09.

\tableofcontents

\parindent 0pt\parskip 0.5ex

\input{session}

\nocite{*}
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
