\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}

\usepackage{authblk}

  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Sorted Rewriting, Conditional Rewriting, and Logically Constrained Rewriting}
\author{Akihisa Yamada}
\affil{National Institute of Advanced Industrial Science and Technology, Japan}

\maketitle

\begin{abstract}
This entry provides various materials for sorted term rewrite systems (sorted TRSs),
(sorted) conditional TRSs (CTRSs),
and
logically constrained TRSs (LCTRSs).
For (C)TRSs we formalize
the fundamental result that the rewrite steps induced by a (C)TRS is the least rewrite relation that models the (C)TRS.
For LCTRSs we simply formulate logics as sorted algebras with the bool sort and logical symbols which are interpreted as expected.
This allows us to define rewrite steps of LCTRSs as rewrite steps of an (infinite) TRS.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introduction}

This entry provides various materials needed for
\emph{logically constrained term rewrite systems} (LCTRS)~\cite{KopN13},
a term rewrite system (TRS) where rules come with constraints
which are interpreted in a background logic.

Building upon the IsaFoR (Isabelle Formalization of Rewriting~\cite{ThiemannS09})
formalizations of unsorted terms \cite{First_Order_Terms-AFP}
and sorted terms \cite{Sorted_Terms-AFP},
we start with formalizing sorted TRSs.
Here we formalize fundamental results such as that the rewrite relation induced by a sorted TRS is the least rewrite relation which models the TRS.
We also formalize (sorted) conditional TRSs (CTRSs) and the fundamental result that
the many-step reduction is the least rewrite preorder which models the CTRS.

As LCTRSs require \emph{logics}, we provide a light-weight and rewriting-oriented formalization of logics.
While it is standard to treat logical symbols and function symbols differently---cf.\@ the IsaFoL (Isabelle Formalization of Logics) entry~\cite{Abstract_Completeness-AFP}---%
in this development we consider a logic just as a sorted algebra where the signature contains logic symbols
and their interpretations are as expected.
This became possible thanks to the formalization of sorted terms, and
is crucial for LCTRSs since the formalism allows logical symbols to appear inside terms.

We also provide means to extend an algebra with another algebra,
and to extend an algebra with logic.
This allows us to build concrete logics modularly:
we define algebras for natural numbers and
integers, and obtain corresponding logics by extending these algebras
with a pure Bool logic.




% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
\endinput
%:%file=~/Projekte/ay/Isabelle_Logic/thys/document/root.tex%:%
