\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

\usepackage{amssymb}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{liiteral}

% for uniform font size
\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{$Pre^*$: The Predecessors of a Regular Language w.r.t.\ a
  Context-Free Grammar, with Applications}
\author{Tassilo Lemke and Tobias Nipkow}
\maketitle

\begin{abstract}
Let $L$ be a language, $G$ a context-free grammar
and let $pre^*(L)$ be the language of all predecessors (w.r.t.\ $G$) of words in $L$.
The following fact has been rediscovered in the literature repeatedly:
If $L$ is regular, so is $pre^*(L)$. Moreover, given an NFA $M$ for $L$, an NFA $M'$ for $pre^*(L)$
can be computed very elegantly from $M$. Starting from a suitable $M$, simple checks on $M'$
provide solutions to many elementary decision problems concerning $G$,
such as the word-problem, emptiness problem, and more.

We formalize two algorithms to compute $pre^*(L)$ for a regular $L$ (using NFAs as representation).
The first one is very simple and elegant and works for any CFG, while the second one is
more efficient but is restricted to CFGs in in extended-CNF.
All our algorithms are executable,
allowing many elementary problems on context-free grammars to be solved automatically.
\end{abstract}

\newpage
\setcounter{tocdepth}{2}
\tableofcontents

\parindent 0pt\parskip 0.5ex

\newpage

\section{Introduction}

Given a regular language $L$ and a context-free grammar $G$, the
language of predecessors of $L$ with respect to $G$, $pre^*(L)$, is
also regular.  This has been discovered independently by many authors
\cite{book1993string,buchi59,caucal1992regular}.  We formalise the
algorithm proposed by Book and Otto \cite{book1993string} which takes
as input a non-deterministic finite automaton $M$, enriches it with
new transitions, and yields a new automaton $M'$ such that
$L(M') = pre^*(L(M))$.

This yields a unified framework for deciding many elementary
properties of context-free grammars, as was first described by Esparza
and Rossmanith \cite{EsparzaR97}.

These theories formalize $pre^*$, its applications to elementary CFG
problems, and an improved algorithm for grammars in CNF by
Bouajjani\emph{ et~al.} \cite{bouajjani2000efficient}.

The theories \textit{Labeled\_Transition\_System} and \textit{LTS\_Automata} are auxiliary;
the formalization proper starts with theory \textit{Pre\_Star}.

Closely related work:
\begin{itemize}
\item \cite{Pushdown_Systems-AFP} formalizes a version of $pre^*$ for
  pushdown systems instead of CFGs.
 \item \cite{Lammich09} formalizes $pre^*$ for dynamic pushdown
   networks, which are a generalization of pushdown systems.
\end{itemize}

\input{session}

\addcontentsline{toc}{section}{References}
\bibliographystyle{alphaurl}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
