\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Nominal Unification}
\author{Guilherme Borges Brand\~ao$^1$ \and Thomas Ammer$^2$ \and Daniele Nantes Sobrinho$^1$ \and Mauricio Ayala-Rinc\'on$^1$ \and  Christian Urban$^2$ \and Maribel Fern\'andez$^2$ \and Mohammad Abdulaziz$^2$\\[2mm]
\mbox{}$^1$ Universidade de Bras\'ilia, Bras\'ilia D.F., Brazil \\[1mm]
\mbox{}$^2$King's College London, London, U.K.
 }

\maketitle

\begin{abstract}
  \noindent
  This is an improved and updated formalisation of the nominal
  unification algorithm.  Nominal unification is a generalisation of
  the classic first-order unification to the practically important
  case of equations between terms involving binding operations. A
  simple, possibly capturing substitution of terms for variables
  solves such equations if it makes the equated terms
  alpha-equivalent, i.e.~equal up to renaming bound names. While
  nominal unification can be seen as equivalent to Miller's
  higher-order pattern unification (unification problems can be inter-coded),
  nominal unification has the advantage of involving only first-order
  syntax. This means in the presented formalisation we only need to
  reason about standard inductive datatypes and first-order
  operations. The main improvement of this development is a much
  simpler proof for establishing the equivalence relation properties
  for an inductive definition characterising when two terms are
  alpha-equivalent. In the original literature, this involved a clunky
  mutual induction over the size of terms involving three properties. Here 
  we can separate the proofs and use just structural inductions. This
  results in a much smoother formalisation of the nominal
  unification algorithm.

\end{abstract}
\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories

\section{Introduction}

This is an improved and updated formalisation of the nominal
unification algorithm given in
\cite{DBLP:journals/corr/abs-1012-4890,DBLP:journals/tcs/UrbanPG04}.
The formalisation presented here is improved according to the Rocha
Oliveira et al.~approach formalised in PVS
\cite{DBLP:journals/entcs/Ayala-RinconFO16,ACRochaOliveira2016}, where
the properties of symmetry, transitivity, and equivariance of the
$\alpha$-equivalence relation are proved separately. This makes the
proof and formalisation more convenient. The same approach was taken
by de Carvalho et al. in the Coq formalisation of nominal
$\alpha$-unification and nominal C-unification
\cite{DBLP:journals/tcs/Ayala-RinconSFN19,WLRdeCarvalho2019}. In the
PVS formalisation, nominal unification is presented as a functional
recursive algorithm, proved sound and complete, from which executable
code in Lisp can be extracted. In the Coq formalisation,
non-deterministic rule-based nominal unification procedures were
presented, following the paper by Urban et al, which were proved sound
and complete; further, recursive definitions were presented, which
were proved equivalent to the inductive procedures, and from which
executable code was extracted.


\input{session}

% optional bibliography
\addcontentsline{toc}{section}{References}
\bibliographystyle{abbrv}
\bibliography{root}



%
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
