\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{pdfsetup}
\urlstyle{rm}
\isabellestyle{it}

\begin{document}
\title{A Reusable Isabelle/HOL Framework for Propositional Labelled Natural Deduction}
\author{Arthur Freitas Ramos \and David Barros Hulak \and Ruy J. G. B. de Queiroz}
\maketitle

\begin{abstract}
A reusable Isabelle/HOL framework presenting labelled natural deduction
as a conservative refinement of ordinary natural deduction, with generic
structural metatheory and several concrete label interpretations. We
define ordinary intuitionistic propositional natural deduction, a
parametric labelled refinement, prove erasure soundness (labelled
derivations are ordinary derivations after erasing labels) and a
controlled lifting theorem (ordinary derivations admit some labelling).
Standard structural metatheory --- weakening, exchange, substitution,
cut admissibility --- is established in both the ordinary and labelled
systems.

The framework is instantiated three times. Unit labels recover ordinary
ND as a special case via conservativity. Provenance labels carry sets
of assumption indices and give both a sound over-approximation of
assumption dependencies and a completeness result equating ordinary
derivability with the existence of a provenance-labelled derivation
whose filtered sub-context still derives the conclusion. A
possible-world modal instance derives the K axiom schema and
necessitation and is connected to Kripke validity; the modal fragment
is presented as an example instantiation, not as a full labelled modal
proof theory.

The development is fully constructive, contains no unfinished proof
commands or oracle invocations, and is intended as a foundation other
entries can cite.
\end{abstract}

\tableofcontents
\newpage

\section{What this entry proves}

The main named results, with the theory in which each is established, are:

\begin{sloppypar}
\begin{itemize}
  \item \texttt{labelled\_sound\_under\_erasure} \\
        (\texttt{Erasure}) --- every labelled derivation erases to an
        ordinary derivation.
  \item \texttt{lifting\_general},
        \texttt{ordinary\_derivations\_admit\_labelling} \\
        (\texttt{Label\_Algebra}, \texttt{Lifting}) --- ordinary
        derivations admit a labelling under the assumed lift-coherence
        law.
  \item \texttt{framework\_adequacy},
        \texttt{framework\_adequacy\_iff} \\
        (\texttt{Label\_Algebra}) --- the locale-level equivalence
        between labelled derivability up to a label and ordinary
        derivability after erasure.
  \item \texttt{unit\_labels\_conservative} \\
        (\texttt{Unit\_Labels}) --- the unit-labels instance recovers
        ordinary ND.
  \item \texttt{provenance\_overapproximates\_dependencies} \\
        (\texttt{Provenance\_Labels}) --- the provenance instance gives
        a sound over-approximation of assumption dependencies.
  \item \texttt{ordinary\_admits\_safe\_provenance}, \\
        \texttt{ordinary\_admits\_tight\_provenance}, \\
        \texttt{provenance\_completeness} \\
        (\texttt{Provenance\_Labels}) --- the provenance instance is
        dependency-tracking complete: ordinary derivability is
        equivalent to the existence of a provenance-labelled derivation
        whose selected filtered context still derives the conclusion.
  \item \texttt{modal\_K\_schema}, \texttt{modal\_necessitation},
        \texttt{modal\_K\_kripke\_valid} \\
        (\texttt{Modal\_K\_Schema}) --- the worked modal instance, where
        the K axiom schema is derived uniformly and connected to Kripke
        validity.
\end{itemize}
\end{sloppypar}

\section{Overview}
Labelled deductive systems were introduced by Gabbay as a general way to attach
structured information to formulae and rules \cite{gabbay-lds}.  The modal
example in this entry follows the standard possible-world reading of modal
labels, closely related to labelled proof-theoretic presentations such as
Simpson's work on intuitionistic modal logic \cite{simpson-modal}. Labelled
approaches to non-classical proof systems have been developed extensively by
Vigan\`o \cite{vigano-labelled-modal} and by Basin, Matthews, and
Vigan\`o \cite{basin-vigano-labelled}; our framework targets the simpler
proof-theoretic core of propositional intuitionistic ND but recovers similar
structural properties.

\section{Label algebra and adequacy}
The entry factors the parametric part of labelled derivability through a
label-algebra locale.  The locale records the proof obligations needed by
any future label instance: that label constructors are uniform under erasure
to the singleton type and that ordinary derivations can be lifted coherently
over the labelled context.  Its central theorem,
\texttt{framework\_adequacy}, states that labelled derivability up to a label
is equivalent to ordinary derivability after erasing the context, with the
chosen lift supplying the canonical labelled witness.  The result is
intentionally architectural rather than surprising: it packages the
soundness/lifting obligations a new label instance must discharge, so that
once those obligations are met the equivalence comes for free.

\section{Modal K axiom and necessitation}
\begin{sloppypar}
The worked modal instance is presented as an example, not as a self-contained
labelled modal proof theory, but it is strong enough to derive the K axiom
schema uniformly in the formulae \(A\), \(B\), the accessibility relation
\(R\), and the world \(w\).  The theory \texttt{Modal\_K\_Schema.thy} records
this as \texttt{modal\_K\_schema}, proves \texttt{modal\_necessitation} under
the standard side-condition that necessitation applies only to formulae
derivable from the empty context at every world and relation, and connects
the labelled derivation to Kripke semantics through
\texttt{modal\_K\_kripke\_valid}.
\end{sloppypar}

\section{Provenance completeness}
The provenance instance is the entry's main dependency-tracking result.
It is adequate in a strong sense: every ordinary derivation lifts to a
labelled derivation whose provenance is contained in the indices of the
ordinary context and whose selected sub-context is still sufficient to
re-derive the conclusion.  The relevant named theorems are:

\begin{sloppypar}
\begin{itemize}
  \item \texttt{ordinary\_admits\_safe\_provenance} ---
        records the index-bound and well-formedness properties of the
        lifted provenance derivation.
  \item \texttt{ordinary\_admits\_tight\_provenance} ---
        strengthens this to a sufficient filtered sub-context.
  \item \texttt{provenance\_completeness} ---
        states the resulting equivalence between ordinary derivability
        and provenance-labelled derivability with a sufficient filtered
        context.
\end{itemize}
\end{sloppypar}

\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
