\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath}
\usepackage{amssymb}

\isabellestyle{it}

\begin{document}

\title{The Busy Beaver Upper-Bound Principle}
\author{Arthur Freitas Ramos\\
David Barros Hulak\\
Ruy J. G. B. de Queiroz}
\maketitle

\begin{abstract}
This development formalizes the Busy Beaver upper-bound principle and its
computability consequences for the Turing machines from AFP's Universal Turing
Machine entry.  The abstract layer works over machine models with finite size
classes and unique exact halting times: it defines the time Busy Beaver function
and proves that any total upper bound for this function decides zero-input
halting.  The development then connects the abstract theory to AFP's Universal
Turing Machine formalization by defining exact numeric-result halting times, a
finite size measure for Turing programs, a concrete Turing-machine Busy Beaver
function, a code-indexed Busy Beaver function whose upper bounds decide AFP's
special halting problem \emph{K1}, and a strengthened program/input-pair Busy
Beaver function whose upper bounds decide AFP's two-argument halting problem
\emph{H1}.  Using AFP's Turing-computability interface, it further proves that
no upper-bound-induced \emph{H1} decider set has a Turing-computable total
characteristic function, with a specialization to the decider induced by the
concrete pair-indexed Busy Beaver function itself.
\end{abstract}

\tableofcontents

\section*{Main Results}

The formalization separates a reusable Busy Beaver upper-bound principle from
its concrete Turing-machine applications.

\begin{itemize}
\item Any total upper bound for the abstract function \emph{BB-time} decides
  zero-input halting for the corresponding model.
\item The concrete function \emph{Turing-BB-time} bounds exact numeric-result
  halting times for bounded-size Turing programs.
\item In AFP's \emph{hpk} locale, code-indexed upper bounds decide the diagonal
  halting problem \emph{K1}.
\item Pair-indexed upper bounds decide the two-argument halting problem
  \emph{H1}, and the induced \emph{H1} characteristic functions are not
  Turing-computable total.
\end{itemize}

\input{session}

\section*{AI Assistance}

AI assistance was used for proof engineering. The final definitions,
statements, and proofs are checked by Isabelle.

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
