\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{amssymb}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Formalization of Lower-Bound Certificates for Optimal Classical Planning}
\author{Dmitriy Traytel}
\maketitle

\begin{abstract}
\looseness=-1
We formalize the framework of pseudo-Boolean lower-bound certificates for
classical planning introduced by Dold, Helmert, Nordstr\"om, R\"oger and
Schindler~\cite{DoldHNRS25}. A lower-bound certificate for a STRIPS (Simplified
Theory of International Planning Representation and Scheduling) planning task
and a bound $B$ is a pseudo-Boolean circuit together with three proofs in
the cutting planes proof system with reification; its validity guarantees that
every plan of the task costs at least $B$, and hence that a plan of cost $B$ is
optimal. We prove the soundness of the certificate format (Theorem~1 of the
paper) and formalize the paper's case study: certificates extracted from a run
of the A$^*$ algorithm, with heuristic certificates for pattern database
heuristics and for the maximum heuristic $h^{\mathit{max}}$, including the
appendix material on efficiently proof-logging pattern databases. The main
results state that a suitable snapshot of a terminated A$^*$ run yields a
valid certificate and therefore proves the optimality of the plan it found.
All locales are shown consistent by concrete interpretations, which also
compose the pattern database certificate with the A$^*$ certificate, as
intended by the paper.
\end{abstract}

\tableofcontents

\section{Introduction}

Optimal classical planners are complex pieces of software, and bugs that lead
to suboptimal plans being reported as optimal are difficult to detect by
testing: while a plan itself is easy to validate, its \emph{optimality} is
not. Dold et al.~\cite{DoldHNRS25} address this problem with proof logging:
the planner emits, alongside the plan, a \emph{lower-bound certificate} that
can be checked by an independent, much simpler verifier in the style of
VeriPB~\cite{BogaertsGMN23}. The certificate asserts that no plan of cost
less than a bound $B$ exists; together with a plan of cost $B$ this
establishes optimality.  (The paper also discusses how a certificate
with a sufficiently large bound~$B$ establishes
unsolvability~\cite{ErikssonRH17}; the formalization here handles
arbitrary finite bounds~$B$.)

This entry formalizes the central trust story of that paper: the
\emph{soundness} of the certificate format, and the existence of valid
certificates for the paper's case study, A$^*$ with pattern
database~\cite{Edelkamp01} and $h^{\mathit{max}}$~\cite{BonetG01} heuristics.
The development is organized as follows.

\begin{description}
\item[\isatext{Lower\_Bound\_Certificates}] formalizes STRIPS planning tasks,
pseudo-Boolean constraints and their $0/1$ semantics, the cutting planes proof
system with reification (CPR), the pseudo-Boolean encoding of a planning task
(Definition~1 of the paper), pseudo-Boolean circuits (Definition~2),
lower-bound certificates (Definition~3), and the soundness theorem
(Theorem~1): a valid certificate for bound $B$ implies that every plan costs
at least $B$.

\item[\isatext{Encoding\_Semantics}] develops the semantic toolkit used by all
later theories. The reverse unit propagation (RUP) rule of the paper is
over-approximated by a semantic rule, so every ``RUP can derive $X$'' lemma of
the paper reduces to a semantic implication over $0/1$ models; the bridge
lemma \isatext{semantic\_to\_cpr} converts such implications into CPR
derivations. The theory also proves characteristic ``forcing'' lemmas for each
constraint of the encoding, and instantiates Theorem~1 at the extended
variable type $\alpha + \gamma$, which provides unboundedly many fresh gate
names for certificate circuits.

\item[\isatext{K\_Gates}] formalizes the clipped cost-threshold gates
$K_{\geq l}$ and the cost lemmas of Appendix~B (monotonicity and step
interaction).

\item[\isatext{Heuristic\_Certificates}] formalizes heuristic certificates
(Definition~4): a heuristic-supplied circuit fragment together with per-state
state, goal, and inductivity lemmas.

\item[\isatext{A\_Star\_Certificates}] formalizes the proof-logging A$^*$
algorithm (equations~(9)--(13); the three CPR proof obligations,
the paper's Lemmas~1--5).  A locale captures the
snapshot of a terminated A$^*$ run---the closed list with $g$-values, the open
list, the expansion-closure property, and a valid heuristic certificate---and
the main theorems show that the assembled circuit is a valid lower-bound
certificate, so any plan costs at least $B$ and a found plan of cost $B$ is
optimal.

\item[\isatext{PDB\_Certificates}] constructs heuristic certificates for
pattern database heuristics (equations~(14)--(16); the three proof
obligations, paper Lemma~6 and Appendix~C) from a
distance table over the abstract state space of a pattern, assuming exactly
the two semantic properties of the table that soundness requires: abstract
goal states have distance $0$, and the distance is consistent along abstract
transitions.

\item[\isatext{Hmax\_Certificates}] constructs heuristic certificates for the
maximum heuristic (equations~(17)--(18); the three proof
obligations of Appendix~D) from the values an
implementation computes anyway, assuming the distilled consequences of the
$h^{\mathit{max}}$ fixed-point equations.

\item[\isatext{Efficient\_PDB}] formalizes the appendix material
(Definition~5; the state-set extension lemma, Lemma~7 of Appendix~E;
the efficient-PDB lemmas, Lemmas~8--18 of Appendix~F): a pattern database circuit
restricted to the abstract states with finite goal distance, with an
additional gate $r_\infty$ covering all remaining abstract states.

\item[\isatext{Locale\_Instances}] interprets every locale of the development
with a concrete planning task, ruling out inconsistent assumption sets. The
interpretations compose the pattern database certificate with the A$^*$
certificate and yield the concrete optimality theorem for the example task.
\end{description}

Two systematic deviations from the paper are worth highlighting; both are
documented where they occur. First, the RUP rule is modeled semantically
(every constraint derivable by unit propagation to conflict satisfies the
semantic condition, so all paper certificates remain expressible), which
means that statements about the \emph{length} of derivations---such as the
$O(|B|)$ bound of Lemma~7, Appendix~E---are out of scope: of such lemmas we formalize the
semantic content only. Second, the $K$-gates referenced by the heuristic
circuits are inlined over the cost bits rather than shared with the search's
own gates, since certificate circuits may only reference input variables and
their own earlier gates. Otherwise, the formalization follows the paper's structure closely;
the mapping from paper lemmas to Isabelle facts is given in the text blocks of each theory.

\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
