\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym,latexsym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

%\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Gr\"obner Bases Theory}
\author{Fabian Immler and Alexander Maletzky\thanks{Supported by the Austrian 
Science Fund (FWF): grant no. W1214-N15, project DK1}}
\maketitle

\begin{abstract}
This formalization is concerned with the theory of Gr\"obner bases in 
(commutative) multivariate polynomial rings over fields, originally developed 
by Buchberger in his 1965 PhD thesis. Apart from the statement and proof of the 
main theorem of the theory, the formalization also implements Buchberger's 
algorithm for actually computing Gr\"obner bases as a tail-recursive function, 
thus allowing to effectively decide ideal membership in finitely generated 
polynomial ideals. Furthermore, all functions can be executed on a concrete 
representation of multivariate polynomials as association lists.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\newpage
\section{Introduction}

The theory of Gr\"obner bases, invented by Buchberger in 
\cite{Buchberger1965,Buchberger1970}, is ubiquitous in many areas of 
computer algebra and beyond, as it allows to effectively solve a multitude of 
interesting, non-trivial problems of polynomial ideal theory. Since its 
invention in the mid-sixties, the theory has already seen a whole range of 
extensions and generalizations, so the present formalization must be understood 
as some kind of ``preliminary'' work that is merely about the most basic 
concepts and results.

For further information about Gr\"obner bases theory one may consult the 
introductory paper \cite{Buchberger1998a} or literally any book on 
commutative/computer algebra, e.\,g. \cite{Adams1994,Kreuzer2000}.

\subsection{Related Work}

The theory of Gr\"obner bases has already been formalized in a couple of other 
proof assistants, listed below in alphabetical order:
\begin{itemize}
 \item ACL2 \cite{Medina-Bulo2010},
 \item Coq \cite{Thery2001,Jorge2009},
 \item Mizar \cite{Schwarzweller2006}, and
 \item Theorema \cite{Buchberger2003,Maletzky2016b}.
\end{itemize}

Please note that this formalization must not be confused with the 
\textit{algebra} proof method based on Gr\"obner bases \cite{Chaieb2007}, which 
is a completely independent piece of work: our results could in principle be 
used to formally prove the correctness and, to some extent, completeness of 
said proof method.

\subsection{Future Work}

This formalization can be extended in several ways:
\begin{itemize}
 \item \emph{Criteria} for avoiding useless reductions in Buchberger's 
algorithm (i.\,e. function \textit{gb}) may be implemented, thus making the 
algorithm much more efficient. See \cite{Buchberger1979} for details.

 \item The degree-reverse-lexicographic term order might be added, 
complementing the existing purely lexicographic and degree-lexicographic term 
orders.

 \item One could establish the connection to \emph{elimination theory}, 
exploiting the well-known \emph{elimination property} of Gr\"obner bases 
w.\,r.\,t. certain term-orders (e.\,g. the purely lexicographic one). This 
would allow one to effectively simplify (and even solve, in some sense) 
systems of algebraic equations.
\end{itemize}

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
