\documentclass[11pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{isabelle,isabellesym}
\usepackage{fontspec}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>


% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}
\title{Axiomatic theory of hereditarily finite sets and its fragments.}
\author{Štěpán Holub \and Zuzana Haniková}
  

 \date{ \today
 \vfill
  Funded by the Czech Science Foundation grant GA\v CR 25-16489S.}
  
\maketitle

\begin{abstract}
Axiomatization of the theory of hereditarily finite sets in classical first-order logic is systematically explored and formalized.
Our list of axioms includes the usual ones of the Zermelo-Fraenkel set theory, as well as some of their less usual variants and alternatives.
The strongest theory considered is ZFfin, obtained from ZF by replacing the axiom of infinity with its negation and adding the axiom of transitive closure.
Of particular interest are the axioms used in the set fragment of the Alternative Set Theory by Vopěnka \cite{Vopenka1979,Sochor1979,Sochor1982,Sochor1983}, which provide a different approach to axiomatizing ZFfin. The hierarchy of fragments of ZFfin is represented as a hierarchy of locales.

Set-theoretic membership, the only basic predicate in the language of the theory, is rendered as a polymorphic binary predicate. Accordingly, first-order formulas of this language are understood as predicates over that type.  ZFfin is not finitely axiomatizable, so it is necessary to use axiom schemata. To that end, an inductive definition of set-theoretically definable predicates is introduced and their basic properties are elaborated.

A major aim of the formalization is to study  several axioms of finiteness and to formalize the equivalence of the fragments in which these axioms are used.
In particular, we formalize the equivalence of ZFfin and the set fragment of Vopěnka's Alternative Set Theory, where the latter uses the schema of induction for sets as a finiteness principle.
Next to that, Tarski finiteness and Dedekind finiteness are used.
The formalization also focuses on the axioms of transitive closure, regularity, and the axiom schema of epsilon induction and their interplay in the context of various fragments.

Hereditarily finite sets, previously formalized by Paulson, satisfy the theory ZFfin.
To demonstrate independence of some statements, several non-standard models of specific fragments are constructed.
We formalize the Beranys-Rieger method of permutation models, and using the construction of a model dating back to \cite{Rieger1957} we formalize the proof that neither the scheme of regularity nor the existence of a transitive closure follow from regularity for finite sets.

This formalization takes inspiration from various sources but does not follow closely any of them.
In addition to already cited literature, it covers some results from  \cite{BaratellaFerro1993}, \cite{Behounek1998}, \cite{ManciniZambella2001} and \cite{KayeWong2007}.
	
	
%Axiomatization of the theory of sets without the axiom of infinity in first-order classical logic is systematically explored and formalized.
%
%We consider standard axioms of the Zermelo-Fraenkel set theory, as well as some less standard ones. In particular, we are interested in axioms that represent the set fragment of the Aletrnative Set Theory of Petr Vopěnka, \cite{Vopenka1979,Sochor1979,Sochor1982,Sochor1983}, in which the key role play the axiom of a set-successor, and the axiom of induction.
%
%An inductive definition of first-order definable predicates is introduced and used to formalize axiom schemata. 
%
%We study several axioms of finiteness and their equivalence in appropriate contexts. In addition to the negation of the standard axiom of infinity, we consider induction, Tarski finiteness and Dedekind finiteness. 
%
%We explore how transitive closures, regularity and well-foundedness behave in different contexts.
%
%We verify that Hereditarily finite sets,as formalized by Paulson, satisfy all our axioms. We also construct several non-standard models of specific fragments, showing independence of some axioms.
%We formalize the method of permutation models, and using the construction of a recursive model \cite{Rieger1957} we formalize the proof that neither the scheme of regularity nor the existence of a transitive closure follow from regularity for finite sets. 
%
%We take inspiration from various sources but we do not follow closely any of them. In addition to already cited literature, our formalization covers some results from \cite{ManciniZambella2002}, \cite{KayeWong2007}, \cite{BaratellaFerro1993} and \cite{Behounek1998}.
\end{abstract}

\tableofcontents
\vspace{\baselineskip}

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{References}
\renewcommand{\bibname}{References}
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
