\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

%\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage[greek,english]{babel}
  %option greek for \<euro>
  %option english (default language) for \<guillemotleft>, \<guillemotright>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Proving the Impossibility of Trisecting an Angle and Doubling the Cube}
\author{Ralph Romanos and Lawrence Paulson}
\maketitle

\begin{abstract}
Squaring the circle, doubling the cube and trisecting an angle, using a compass and 
straightedge alone, are classic unsolved problems first posed by the ancient Greeks. 
All three problems were proved to be impossible in the 19th century. The following document 
presents the proof of the impossibility of solving the latter two problems using Isabelle/HOL, 
following a proof by Carrega. The proof uses elementary methods: no Galois theory or field 
extensions. 
The set of points constructible using a compass and straightedge is defined inductively. 
Radical expressions, which involve only square roots and arithmetic of rational numbers, 
are defined, and we find that all constructive points have radical coordinates. Finally, 
doubling the cube and trisecting certain angles requires solving certain cubic equations 
that can be proved to have no rational roots. The Isabelle proofs require a great many 
detailed calculations.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
%\bibliographystyle{abbrv}
%\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
