\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{amssymb}
\usepackage{pdfsetup}

\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{First Welfare Theorem \thanks{%
This work is supported by the Austrian Science Fund (FWF) project P26201 and the European Research Council (ERC) grant no 714034 \emph{SMART}.}}

\author{Julian Parsert \and Cezary Kaliszyk}

\maketitle

\begin{abstract}
  Economic activity has always been a fundamental part of society.
  Due to modern day politics, economic theory has gained even more
  influence on our lives. Thus we want models and theories to be as
  precise as possible. This can be achieved using
  certification with the help of formal proof technology. Hence we
  will use Isabelle/HOL to construct two economic models, that of the the pure
  exchange economy and a version of the Arrow-Debreu Model. We will
  prove that the \emph{First Theorem of Welfare Economics}, holds
  within both. The theorem is the mathematical formulation of Adam
  Smith's famous \emph{invisible hand} and states that a group of
  self-interested and rational actors will eventually achieve an
  efficient allocation of goods and services.
\end{abstract}

\tableofcontents

\parindent 0pt\parskip 0.5ex
\input{session}

\section{Related work}
%
There have been multiple attempts at formalizing economical concepts.
Most notably the ForMaRE project~\cite{DBLP:conf/mkm/0002RK13}. The
project intended to apply formal mathematical reasoning to economic
results. One way of doing so, was the development of the ``Auction
Theory Toolbox''~\cite{kerber2013developing}. This entailed the
formalization of auction mechanisms such as the Combinatorial Vickrey
auctions~\cite{DBLP:journals:corr:CaminatiK0R13}. Having proved the
auction mechanisms' soundness among other properties, the team used
Isabelle to generate executable code, which conforms with that formal
specification~\cite{DBLP:journals:corr:CaminatiK0R13}.
%
Interestingly, there have been very few formalizations of results from
\emph{Game Theory}. One of the few examples is St{\'{e}}phane Le
Roux's~\cite{DBLP:conf/tphol/Roux09} formalization of \emph{Nash
  Equilibria} in Coq. In \emph{Social Choice Theory}, another
mathematical area with applications to economics, Freek
Wiedijk~\cite{Wiedijk2009} formalized Arrow's famous
\emph{Impossibility Theorem} using Mizar. The same theorem was also
formalized by Tobias Nipkow in
Isabelle~\cite{DBLP:journals/afp/Nipkow08b}. Furthermore, Peter
Gammie~\cite{SenSocialChoice-AFP} formalized \emph{May's Theorem} and
\emph{Sen's Theorem} as well as other \emph{Social Choice Theory}
results. More recently, Manuel
Eberl~\cite{Randomised_Social_Choice-AFP} formalized multiple basics
of \emph{Randomised Social Choice Theory}. Finally, aspects of finance
have been formalized by Peter J{\"a}ger~\cite{PeterJaeger2012} as well as 
Echenim and Peltier~\cite{DBLP:conf/cade/EchenimP17}


\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
