\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}

\usepackage{authblk}

  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Verifying a Decision Procedure for Pattern Completeness\footnote{This research was supported by the Austrian Science Fund (FWF) project I~5943.}}
\author{Ren\'e Thiemann}
\affil{University of Innsbruck, Austria}
\author{Akihisa Yamada}
\affil{National Institute of Advanced Industrial Science and Technology, Japan}
\maketitle

\begin{abstract}
Pattern completeness is the property that the left-hand sides of a functional
program or term rewrite system cover all cases w.r.t.\ pattern matching. 
We verify a recent (abstract) decision procedure 
for pattern completeness that covers the general case, i.e., in particular 
without the usual restriction of left-linearity.
In two refinement steps, we further develop an 
executable version of that abstract algorithm.
On our example suite, this verified implementation is faster 
than other implementations that are based
on alternative (unverified) approaches, including the complement algorithm, 
tree automata
encodings, and even the pattern completeness check of the GHC Haskell compiler.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introduction}

This AFP entry includes the formalization of a decision procedure \cite{FSCD2024} 
for pattern completeness. It also contains 
the setup for running the experiments of that paper, i.e., 
it contains
\begin{itemize}
\item a generator for example term rewrite systems and Haskell programs of varying size,
\item a connection to an implementation of the complement algorithm \cite{LazrekLT90}
  within the ground confluence prover AGCP~\cite{AotoT2016}, and
\item a tree automata encoder of pattern completeness that is linked 
  with the tree automata library FORT-h~\cite{FORTh}.
\end{itemize}

Note that some further glue code is required to run the experiments, which is not
included in this submission. Here, we just include the glue code that was defined
within Isabelle theories.


% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
\endinput
%:%file=~/Projekte/ay/Isabelle_Logic/thys/document/root.tex%:%
