\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

\usepackage{latexsym}
\usepackage{wasysym}

%\usepackage{eurosym}
  %for \<euro>

\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Conditional normative reasoning as a fragment of HOL
  (Isabelle/HOL dataset)}
\author{Xavier Parent and Christoph Benzm{\"u}ller}
\maketitle

\begin{abstract}
 We present a mechanisation of (preference-based) conditional
 normative reasoning. Our focus is on \AA qvist's system {\bf E} for
 conditional obligation and its extensions. We present both a
 correspondence-theory-focused metalogical study and a use-case
 application to Parfit's repugnant conclusion, focusing on the mere addition paradox.
Our contribution is explained in detail in \cite{J68}.  This document
presents a corresponding (but sligthly modified) Isabelle/HOL dataset.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex


\section{Introduction}
In this document we present the Isabelle/HOL dataset associated with
\cite{J68}, in which 
%we describe our mechanization of  \AA qvist's system {\bf E} for
%conditional obligation, and its extensions. We illustrate two key uses of the framework: first, for meta-reasoning about the logic, particularly for verifying %deontic correspondences similar to modal logic; second, for assessing ethical arguments, exemplified by encoding Parfit's mere addition paradox in %population ethics, a smaller version of his so-called repugnant conclusion.
%(preference-based) conditional
%normative reasoning. Our focus is on
\textit{``We report on the mechanization of (preference-based) conditional
normative reasoning. Our focus is on \AA qvist's system {\bf E} for
conditional obligation, and its extensions. Our mechanization is
achieved via a shallow semantical embedding in Isabelle/HOL.  We
consider two possible uses of the framework. The first one is as a
tool for meta-reasoning about the considered logic. We employ it for
the automated verification of deontic correspondences  (broadly
conceived) and related matters, analogous to what has been previously
achieved for the modal logic cube.  The equivalence is automatically
verified in one direction, leading from the property to the axiom. The
second use is as a tool for assessing ethical arguments. We provide a
computer encoding of a well-known paradox (or impossibility theorem)
in population ethics, Parfit's repugnant conclusion.''} \cite{J68}
% generated text of all theories
\input{session}

\section{Conclusion}

In this document we presented the Isabelle/HOL dataset associated with
\cite{J68}. We described our shallow semantic embedding of \AA qvist's dyadic deontic logic  {\bf E} and its extensions. We showcased two key uses of the framework: first, for meta-reasoning about the logic, particularly for verifying deontic correspondences similar to modal logic; second, for assessing ethical arguments, exemplified by encoding Parfit's mere addition paradox, a smaller version of his so-called repugnant conclusion.


\section*{Funding Statement} X. Parent's research was funded in whole or in part by the Austrian Science Fund (FWF) [10.55776/M3240, ANCoR project]. For open access purposes, the author has applied a CC BY public copyright license to any author-accepted manuscript version arising from this submission. The work was also supported by the FWF project 10.55776/I6372.
% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
