\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

%\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage[greek,english]{babel}
  %option greek for \<euro>
  %option english (default language) for \<guillemotleft>, \<guillemotright>

%\usepackage[latin1]{inputenc}
  %for \<onesuperior>, \<onequarter>, \<twosuperior>, \<onehalf>,
  %\<threesuperior>, \<threequarters>, \<degree>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<cent>, \<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{Locally Nameless Sigma Calculus}
\author{Ludovic Henrio and Florian Kamm\"uller and Bianca Lutz and Henry Sudhof}
\maketitle

\begin{abstract}
We present a Theory of Objects based on the original functional
$\varsigma$-calculus by Abadi and Cardelli \cite{AC96a} but with an
additional parameter to methods. We prove confluence of the operational
semantics following the outline of Nipkow's proof of confluence for the
$\lambda$-calculus reusing his general \texttt{Commutation.thy}
\cite{nip:01} a generic diamond lemma reduction. 
We furthermore formalize a simple type system for our $\varsigma$-calculus
including a proof of type safety.
The entire development uses the concept of Locally Nameless representation
for binders \cite{ACPPW:POPL08}. The first part, i.e. $\varsigma$-calculus
confluence, has been published on an earlier version based on lists and
deBruijn indices \cite{HK:FMOODS07}.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
%\bibliographystyle{abbrv}
%\bibliography{root}

\bibliographystyle{entcs}
\begin{thebibliography}{10}
\bibitem{AC96a}
M. Abadi and L. Cardelli.
\newblock {``A Theory of Objects''}.
\newblock Springer, New York, 1996.


\bibitem{ACPPW:POPL08}
B. Aydemir, A, Charguéraud, B.~C. Pierce, R. Pollack, and
  S. Weirich.
\newblock Engineering formal metatheory.
\newblock {\em Princ. of Programming Languages, POPL'08}, ACM, 2008. 

\bibitem{HK:FMOODS07}
L. Henrio and F. Kammüller.
\newblock A mechanized model of the theory of objects.
\newblock {\em Formal Methods for Open Object-Based Distributed Systems,}. 
LNCS \textbf{4468} Springer, 2007.


\bibitem{nip:01}
Tobias Nipkow.
\newblock{More Church Rosser Proofs.}
\newblock{\em Journal of Automated Reasoning}. 
\newblock{\textbf{26}:51--66, 2001.}
\end{thebibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
