\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage[a4paper,margin=2.4cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{microtype}
\usepackage{url}

\urlstyle{rm}
\isabellestyle{it}
\setlength{\emergencystretch}{3em}

\begin{document}

\title{Alon's Combinatorial Nullstellensatz}
\author{Arthur F. Ramos \and David Barros Hulak \and Ruy J. G. B. de Queiroz}
\maketitle

\begin{abstract}
This entry formalizes Alon's Combinatorial Nullstellensatz for sparse
multivariate polynomials over fields. The proof derives the coefficient formula
from univariate Lagrange interpolation and uses it to obtain the standard
nonvanishing conclusion over finite grids. AI assistance was used for proof
engineering. The final definitions, statements, and proofs are checked by
Isabelle.
\end{abstract}

\section{Overview}

Alon's Combinatorial Nullstellensatz is a polynomial-method theorem about a
distinguished coefficient of a multivariate polynomial. In the form formalized
here, if all monomials have total degree bounded by $d_1+\cdots+d_n$ and the
coefficient of $X_1^{d_1}\cdots X_n^{d_n}$ is nonzero, then the polynomial has a
nonzero value on every product grid whose $i$-th side has more than $d_i$
elements.

The Isabelle/HOL development represents multivariate polynomials sparsely, as
finite-support coefficient functions from exponent lists to field elements. It
first proves the necessary univariate Lagrange interpolation identities, then
sums polynomial evaluations over a finite grid with the corresponding Lagrange
weights. All monomials except the distinguished exponent vector cancel, yielding
the coefficient formula and the nonvanishing theorem.

\section{Sources}

The entry follows Alon's original paper~\cite{Alon1999}. The formal proof uses
only Isabelle/HOL's standard computational algebra library as its mathematical
base.

\tableofcontents

\sloppy
\input{session}
\fussy

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
