\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}

\usepackage{authblk}

  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Sorted Terms\footnote{This research was partly supported by the Austrian Science Fund (FWF) project I~5943.}}
\author{Akihisa Yamada}
\affil{National Institute of Advanced Industrial Science and Technology, Japan}
\author{Ren\'e Thiemann}
\affil{University of Innsbruck, Austria}
\maketitle

\begin{abstract}
This entry provides a basic library for many-sorted terms and algebras.
We view sorted sets just as partial maps from elements to sorts,
and define sorted set of terms reusing the data type
from the existing library of (unsorted) first order terms.
All the existing functionality, such as substitutions and contexts,
can be reused without any modifications.
We provide predicates stating what substitutions or contexts are considered sorted,
and prove facts that they preserve sorts as expected.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introduction}

This entry extends the First-Order Terms \cite{First_Order_Terms-AFP}
entry with many-sorted terms.
Instead of defining a new datatype for sorted terms,
we just define sorted sets over the existing datatype of unsorted terms.
We do not even introduce our type for sorted sets:
we just view sorted sets as partial maps from elements to their sorts.

Part of the entry is presented in \cite{FSCD2024}.



% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
\endinput
%:%file=~/Projekte/ay/Isabelle_Logic/thys/document/root.tex%:%
