\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath}
\usepackage{amssymb}

\isabellestyle{it}

\begin{document}

\title{Andrew's Monotone Chain Convex Hull Algorithm}
\author{Arthur Freitas Ramos\\
David Barros Hulak\\
Ruy J. G. B. de Queiroz}
\maketitle

\begin{abstract}
This development formalizes the executable core of Andrew's monotone chain
convex hull algorithm~\cite{Andrew1979}.  The algorithm sorts planar points
lexicographically, removes duplicates, computes lower and upper hull chains by
a stack scan that removes non-left turns, and concatenates the two chains with
their repeated endpoints removed.  The formalization proves the stack-scan
turn invariant, subset and distinctness properties for the computed chains,
ordered-chain facts for the lower and upper scans, a distinctness criterion for
the final concatenation, length bounds, support-function invariants for the
lower and upper scans, and the real-coordinate theorem that the computed output
has the same convex hull as the input.  The final correctness theorem states
the specification explicitly: the returned vertices are input points, every
input point lies in the convex hull of the returned vertices, and the returned
vertices have exactly the same convex hull as the input.  It also proves
irredundancy of the returned vertex set: deleting any returned point changes
the convex hull of the returned set.  The executable algorithm is polymorphic
over ordered integral domains because the scan only uses lexicographic ordering
and signs of orientation determinants; the geometric theorem is stated for real
coordinates because the convexity and separating-hyperplane libraries use real
vector spaces.  Thus the integer examples exercise the executable code, while
the real examples instantiate the geometric correctness theorem.  The
support-function argument shows that points removed by a scan are dominated in
the relevant support directions; a separating hyperplane theorem then yields
convex-hull coverage, and strict supporting directions expose each returned
vertex for the irredundancy result.  The
development proves functional correctness, but does not formalize the
asymptotic running time.  AI assistance was used for proof engineering.  The
final definitions, statements, and proofs are checked by Isabelle.
\end{abstract}

\tableofcontents

\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
