\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}

\isabellestyle{it}

\begin{document}

\title{Nash Equilibria for Finite Games in Isabelle/HOL}
\author{Arthur Freitas Ramos \and David Barros Hulak \and Ruy J. G. B. de Queiroz}
\maketitle

\begin{abstract}
This development formalizes Nash equilibria for finite strategic-form games,
following Nash's equilibrium concept~\cite{Nash1950,Nash1951}.  It gives
reusable definitions of profiles, unilateral deviations, best responses,
dominant strategies, and pure Nash equilibria; proves existence for finite
ordinal potential games~\cite{MondererShapley1996} and games with dominant
strategies; and verifies matching pennies as a finite game with no pure Nash
equilibrium.  It also formalizes mixed-strategy profiles for finite games,
support lemmas for equilibrium strategies, Dirac embeddings of pure profiles,
and the existence of a mixed Nash equilibrium using Brouwer's fixed point
theorem.  Worked examples cover the Prisoner's Dilemma, a coordination game,
matching pennies, and rock-paper-scissors.  AI assistance was used for proof
engineering.  The final definitions, statements, and proofs are checked by
Isabelle.
\end{abstract}

\section*{Contributions and Scope}

This entry gives a reusable Isabelle/HOL formalization of pure and mixed Nash
equilibria for finite strategic-form games.  The pure-game locale supports a
finite player set with player-indexed finite strategy sets and is used to
formalize unilateral deviations, best responses, dominant strategies, and
ordinal potential games.  The mixed-game development proves Nash's finite-game
existence theorem by defining an excess-payoff map on a compact convex product
of simplices and applying Brouwer's fixed point theorem from HOL-Analysis.

The mixed-game locale is intentionally less general than the pure-game locale:
players and pure strategies are represented by finite HOL types, so every
player uses the same finite pure-strategy type.  This choice makes mixed
profiles Cartesian vectors indexed by player/strategy pairs, which gives direct
access to the compactness, convexity, continuity, and fixed-point infrastructure
needed for the Brouwer proof.  The entry also includes support lemmas for mixed
equilibria, Dirac embeddings of pure profiles, and checked examples of the
Prisoner's Dilemma, a coordination game, matching pennies, and
rock-paper-scissors.

\section*{Related Work}

Le Roux, Martin-Dorel, and Smaus formalized a Nash-equilibrium existence theorem
in both Coq and Isabelle for finite-outcome games derived from win/lose
games~\cite{LeRouxMartinDorelSmaus2017}.  Bagnall, Merten, and Stewart
developed an Ssreflect/Coq library for algorithmic game theory, including pure
and mixed Nash equilibria, potential games, smooth games, approximate
equilibria, and applications to routing and congestion games~\cite{BagnallMertenStewart2017}.
The present entry is narrower in mathematical scope but focuses on an
AFP-style Isabelle/HOL development for finite strategic-form games, with a
Brouwer-based mixed-equilibrium existence proof and small canonical examples.

\tableofcontents
\newpage

\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
