\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>
\usepackage{amsmath}

%\usepackage{eurosym}
  %for \<euro>

\usepackage[only,bigsqcap]{stmaryrd}
%for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{The L\'evy-Prokhorov Metric}
\author{Michikazu Hirata}
\maketitle
\begin{abstract}
  We formalize the L\'evy-Prokhorov metric, a metric on finite measures,
  mainly following the lecture notes by Gaans~\cite{Onno}.
  This entry includes the following formalization.
  \begin{itemize}
  \item Characterizations of closed sets, open sets, and topology by limit.
  \item A special case of Alaoglu's theorem.
  \item Weak convergence and the Portmanteau theorem.
  \item The L\'evy-Prokhorov metric and its completeness and separability.
  \item The equivalence of the topology of weak convergence and
  the topology generated by the L\'evy-Prokhorov metric.
  \item Prokhorov's theorem.
  \item Equality of two $\sigma$-algebras on the space of finite measures.
  One is the Borel algebra of the L\'evy-Prokhorov metric and the other is the least $\sigma$-algebra
  that makes $(\lambda\mu.\: \mu(A))$ measurable for all measurable sets $A$.
  \item The space of finite measures on a standard Borel space is also a standard Borel space.
  \end{itemize}
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
