\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pdfsetup}

\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Checking Equality-Saturation Merge and Extraction Certificates}
\author{Arthur Freitas Ramos \and David Barros Hulak \and Ruy J.\ G.\ B.\ de Queiroz}
\maketitle

\begin{abstract}
Equality-saturation engines maintain equalities in an e-graph and later extract
a representative from an e-class.  This entry verifies an executable checker
for the corresponding engine boundary.  It directly parses the flat annotated
S-expression format returned by egg 0.11.0's explanation API
~\cite{egg,egg-api}; flat explanations may apply numbered rewrite rules at
positions or reuse equalities from earlier checked e-class merges.  For
extraction, a finite e-graph is represented as a topologically ordered e-class
DAG.  An executable bottom-up dynamic program selects a representative, and a
formal proof establishes minimum additive cost over every term represented by
the designated class.  Class-aligned equality certificates additionally prove
the selected term equivalent to the canonical class term.

The development reuses the term, substitution, position, context, and rewriting
infrastructure of the AFP entries \emph{First-Order Terms} and
\emph{First-Order Rewriting}.  Its soundness statements target their existing
conversion relation.  It does not introduce another first-order term datatype,
rewrite relation, generic equational proof calculus, or proof of Birkhoff's
theorem.  Those generic equational results and a checker for equality proofs
already exist in IsaFoR/CeTA.  The new material is the chronological e-graph
merge discipline, positional reuse of recorded merges, certified e-class DAG
semantics, and globally optimal dynamic-programming extraction.
\end{abstract}

\tableofcontents
\parindent 0pt\parskip 0.5ex

\section{Reused foundations}

This entry is an extension of existing rewriting libraries, not an alternative
foundation.  It imports the AFP sessions \texttt{First\_Order\_Terms} and
\texttt{First\_Order\_Rewriting}~\cite{afp-terms,afp-rewriting}.  In particular,
it directly uses:
\begin{itemize}
  \item the AFP datatype \texttt{term}, substitutions and substitution
        application;
  \item AFP positions, valid-position sets, subterms, contexts,
        \texttt{replace\_at}, and executable position enumeration; and
  \item AFP rewrite rules, \texttt{rstep}, and the symmetric reflexive-transitive
        closure used for conversions.
\end{itemize}

IsaFoR's equational-reasoning theory already defines the \texttt{eq\_proof}
datatype with reflexivity, symmetry, transitivity, assumption, and congruence
constructors; its executable \texttt{proves} function checks such proofs; and
the same theory relates provability, conversions, semantic entailment, and
Birkhoff completeness
~\cite{isafor-equational-reasoning}.  CeTA also certifies equality proofs and
completion certificates
~\cite{recording-completion,certified-equational-reasoning,isafor-ceta}.  We
therefore do not reproduce any of those definitions or results.

The AFP entry \emph{Proof Terms for Term Rewriting} formalizes proof terms and
their correspondence with multisteps~\cite{afp-proof-terms}.  Those proof terms
represent reductions; the certificates here instead expose chronological
e-class sharing and justify extraction over a finite represented DAG.

\section{Difference from IsaFoR/CeTA}

The generic equational content of a rule-only explanation is intentionally
delegated to the AFP/IsaFoR foundation.  The contribution here begins where an
e-graph exposes data that a generic equational proof does not record:
\begin{enumerate}
  \item The egg importer parses the public \texttt{get\_flat\_strings} format,
        locates its nested \texttt{Rewrite=>}/\texttt{Rewrite<=} annotation,
        resolves the named rule, and infers the substitution with the AFP
        matcher.  The compiled step is then replayed independently.
  \item A \texttt{Merge\_At} step cites a concrete equality by its index in the
        e-graph's merge history and reuses it at an AFP position.
  \item A merge log is checked chronologically.  Entry $i$ may cite only entries
        before $i$, which rules out circular e-class justifications.
  \item A finite acyclic e-graph is represented by nonempty e-classes in
        topological order.  Its semantics contains every e-node and every
        recursive combination of represented child terms.
  \item An executable bottom-up dynamic program memoizes one optimum per class.
        The selected term is proved minimum-cost among all terms represented by
        the designated e-class.
  \item Class-aligned flat certificates prove the canonical e-node equal to
        every alternative e-node.  Congruence then proves every recursively
        represented term equal to the canonical class term.
  \item A separate candidate-set checker supports interchange formats that
        enumerate terms explicitly.
  \item A bounded producer accepts an arbitrary, untrusted step generator, but
        adds only steps accepted by the independent checker.  Every emitted
        path is therefore a valid AFP conversion.
\end{enumerate}

Thus this checker is not a replacement for CeTA's equational reasoning
checker.  It is an equality-saturation-specific front end whose accepted
claims are reduced to the existing AFP conversion relation.

\section{Fixture generated by egg}

The end-to-end fixture uses egg 0.11.0 with its \texttt{deterministic} feature.
The crates.io checksum is:
\begin{center}
{\scriptsize\texttt{dd40cfd4196d7a8f882ace95d623d4d6734588e502b4e188675a7c2f55eb4fb4}}
\end{center}
The generator enabled explanations before adding the seed expression, disabled
explanation-length optimization, ran the four named pattern rules shown in
\texttt{Examples\_Egg}, and executed:
\begin{verbatim}
# Cargo.toml
egg = { version = "=0.11.0", features = ["deterministic"] }

// Rust
let mut explanation =
    runner.explain_equivalence(&left, &right);
explanation.check_proof(&rules);
for step in explanation.get_flat_strings() {
    println!("{step}");
}
\end{verbatim}
Repeated generator runs produced byte-identical output.  The forward fixture
consumed by Isabelle is:
\begin{verbatim}
(+ (shl x 1) (* y 2))
(+ (Rewrite=> shl-one (+ x x)) (* y 2))
(+ (+ x x) (Rewrite=> mul-two (+ y y)))
\end{verbatim}
The backward fixture is:
\begin{verbatim}
z
(Rewrite<= mul-one (* z 1))
\end{verbatim}
An additional regression fixture,
\texttt{(f y (Rewrite=> add-zero z))}, checks that a rewrite in the second
child is assigned AFP position $[1]$ even when the first child is an atom.
These strings are parsed during evaluation of the Isabelle checker; they are
not manually replaced by preconstructed certificate values.

\section{Verified interface}

The main results are:
\begin{itemize}
  \item \texttt{check\_egg\_explanation\_sound}: an accepted textual flat
        explanation emitted by egg denotes an AFP conversion.  The checked
        fixtures exercise both forward and backward annotations inside actual
        egg output.
  \item \texttt{check\_explanation\_sound}: an accepted flat explanation,
        relative to sound recorded merges, is an AFP conversion.
  \item \texttt{checked\_merge\_log\_sound}: every equality in an accepted
        chronological merge log is an AFP conversion generated by the input
        rules.
  \item \texttt{egraph\_explanation\_sound}: a later explanation may safely
        reuse any entry of an accepted merge log.
  \item \texttt{represented\_eclass\_finite}: every class of a well-formed
        acyclic e-graph represents a finite set of terms.
  \item \texttt{extract\_eclass\_global\_minimum}: the executable dynamic
        program returns a represented term no more expensive than any term
        represented by the designated class.
  \item \texttt{checked\_egraph\_eclass\_sound}: every pair of terms
        represented by the same checked class is related by an AFP conversion.
  \item {\footnotesize\texttt{certified\_egraph\_extraction\_correct}}: checked
        equality, representedness, and global minimum cost hold simultaneously
        for the extracted term.
  \item \texttt{run\_bounded\_search\_accepted} and
        \texttt{find\_explanation\_sound}: every path returned by the bounded
        producer is accepted by the checker and denotes an AFP conversion,
        without assumptions on the proposal generator.
\end{itemize}

The acyclic-DAG example has a child class with three equivalent
strength-reduction forms and a root e-node that references that class twice.
The root therefore represents all nine child combinations.  The checked
dynamic program selects the globally cheapest one and proves it equal to the
canonical root.  The executable functions are exported to SML and Haskell.
Another example embeds literal output generated by egg 0.11.0 with deterministic
iteration and explanation-length optimization disabled.  egg's own
\texttt{check\_proof} accepted the traces; the Isabelle checker then parses and
independently replays them against AFP rewriting.

\section{Scope}

No union-find implementation, congruence-closure algorithm, rebuilding,
e-matching, scheduling policy, or e-graph construction algorithm is verified.
The global minimum ranges over all terms represented by the supplied finite
acyclic DAG; it does not claim that an untrusted engine supplied every e-node
it discovered.  A separate candidate-set checker has only list-relative
minimality.  The DAG is ground, with source-language atoms represented as
nullary symbols, as in egg's recursive expressions.  The textual importer
currently supports egg S-expressions with unquoted atoms and pattern-based
named rules; custom appliers and quoted rule names require an adapter before
checking.

\input{session}

\nocite{egg}
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
