\documentclass[11pt,a4paper]{report}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

\usepackage{wasysym}
\usepackage{pifont}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Abstract Consistency Properties}
\author{Asta Halkjær Boserup, Anders Schlichtkrull}
\maketitle

\abstract{
  Smullyan used abstract consistency properties to great effect in unifying meta-theoretical results in first-order logic.
  By abstractly specifying when a set of formulas is consistent, he proved a single model existence result for all sets that met the specification, reusing this result in proving completeness, compactness, etc.
  Fitting later defined abstract consistency properties for a range of other logics.

  In this work we use locales to mechanize abstract consistency properties without fixing a particular logic or syntax, generalizing the work of Smullyan and Fitting.
  We use Fitting's technique for closing a consistency property under limits to guarantee the existence of a maximal element.
  This yields a maximal consistent set for any notion of consistency expressible in the framework.
  The usual conjunctive, disjunctive, universal and existential conditions of abstract consistency properties ---based on Smullyan's uniform notation--- arise as special cases of our abstract development.
  Users of the framework can define an abstract consistency property for their own syntax and logic, prove that it is well behaved, and receive a corresponding maximal consistent set from which to prove model existence.

  We provide three example instantiations.
  First, compactness and completeness for a first-order logic where we only instantiate universal quantifiers with already occurring terms.
  Second, completeness over general models for a second-order logic.
  Third, completeness of our own natural deduction system for Prior's Ideal Language, a recently developed hybrid logic with propositional quantification.
}

\nocite{*}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
