\documentclass[fontsize=11pt,paper=a4,open=right,twoside,abstract=true]{scrreprt}
\usepackage{fixltx2e}
\usepackage{isabelle,isabellesym}
\usepackage[nocolortable, noaclist]{hol-ocl-isar}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[numbers, sort&compress, sectionbib]{natbib}
\usepackage[caption=false]{subfig}
\usepackage{lstisar}
\usepackage{tabu}
\usepackage[]{mathtools}
\usepackage{prooftree}
\usepackage[english]{babel}
\usepackage[pdfpagelabels, pageanchor=false, plainpages=false]{hyperref}
% \usepackage[draft]{fixme}

% MathOCl expressions
\colorlet{MathOclColor}{Black}
\colorlet{HolOclColor}{Black}
\colorlet{OclColor}{Black}
%
\sloppy 

\uchyph=0
\graphicspath{{data/},{figures/}}
\allowdisplaybreaks

\renewcommand{\HolTrue}{\mathrm{true}}
\renewcommand{\HolFalse}{\mathrm{false}}
\newcommand{\ptmi}[1]{\using{\mi{#1}}}
\newcommand{\Lemma}[1]{{\color{BrickRed}%
    \mathbf{\operatorname{lemma}}}~\text{#1:}\quad}
\newcommand{\done}{{\color{OliveGreen}\operatorname{done}}}
\newcommand{\apply}[1]{{\holoclthykeywordstyle%
    \operatorname{apply}}(\text{#1})}
\newcommand{\fun} {{\holoclthykeywordstyle\operatorname{fun}}}
\newcommand{\definitionS} {{\holoclthykeywordstyle\operatorname{definition}}}
\newcommand{\where} {{\holoclthykeywordstyle\operatorname{where}}}
\newcommand{\datatype} {{\holoclthykeywordstyle\operatorname{datatype}}}
\newcommand{\types} {{\holoclthykeywordstyle\operatorname{types}}}
\newcommand{\pglabel}[1]{\text{#1}}
\renewcommand{\isasymOclUndefined}{\ensuremath{\mathtt{invalid}}}
\newcommand{\isasymOclNull}{\ensuremath{\mathtt{null}}}
\newcommand{\isasymOclInvalid}{\isasymOclUndefined}
\DeclareMathOperator{\inv}{inv}
\newcommand{\Null}[1]{{\ensuremath{\mathtt{null}_\text{{#1}}}}}
\newcommand{\testgen}{HOL-TestGen\xspace}
\newcommand{\HolOption}{\mathrm{option}}
\newcommand{\ran}{\mathrm{ran}}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\typedef}{\mathrm{typedef}}
\newcommand{\mi}[1]{\,\text{#1}}
\newcommand{\state}[1]{\ifthenelse{\equal{}{#1}}%
  {\operatorname{state}}%
  {\operatorname{\mathit{state}}(#1)}%
}
\newcommand{\mocl}[1]{\text{\inlineocl|#1|}}
\DeclareMathOperator{\TCnull}{null}
\DeclareMathOperator{\HolNull}{null}
\DeclareMathOperator{\HolBot}{bot}


% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}
\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.\xspace}
\renewcommand{\isamarkupheader}[1]{\chapter{#1}}
\renewcommand{\isamarkupsection}[1]{\section{#1}}
\renewcommand{\isamarkupsubsection}[1]{\subsection{#1}}
\renewcommand{\isamarkupsubsubsection}[1]{\subsubsection{#1}}
\renewcommand{\isamarkupsect}[1]{\section{#1}}
\renewcommand{\isamarkupsubsect}[1]{\susubsection{#1}}
\renewcommand{\isamarkupsubsubsect}[1]{\subsubsection{#1}}

\begin{document}
\renewcommand{\subsubsectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section}
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\chapterautorefname}{Chapter}
\newcommand{\subtableautorefname}{\tableautorefname}
\newcommand{\subfigureautorefname}{\figureautorefname}

\title{Featherweight OCL}
\subtitle{A Proposal for a Machine-Checked Formal Semantics for OCL 2.5}
\author{%
  \href{http://www.brucker.ch/}{Achim D. Brucker}\footnotemark[1]
  \and
  \href{https://www.lri.fr/~tuong/}{Fr\'ed\'eric Tuong}\footnotemark[3]
  \and
  \href{https://www.lri.fr/~wolff/}{Burkhart Wolff}\footnotemark[2]}
\publishers{%
  \footnotemark[1]~SAP AG, Vincenz-Priessnitz-Str. 1, 76131 Karlsruhe,
  Germany \texorpdfstring{\\}{} \href{mailto:"Achim D. Brucker"
    <achim.brucker@sap.com>}{achim.brucker@sap.com}\\[2em]
  %
  \footnotemark[3]~Univ. Paris-Sud, IRT SystemX, 8 av.~de la Vauve, \\
  91120 Palaiseau, France\\
  frederic.tuong@\{u-psud, irt-systemx\}.fr\\[2em]
  %
  \footnotemark[2]~Univ. Paris-Sud, Laboratoire LRI, UMR8623, 91405 Orsay, France\\
  CNRS, 91405 Orsay, France\texorpdfstring{\\}{}
  \href{mailto:"Burkhart Wolff" <burkhart.wolff@lri.fr>}{burkhart.wolff@lri.fr}
}


\maketitle

\begin{abstract}
  The Unified Modeling Language (UML) is one of the few modeling
  languages that is widely used in industry. While UML is mostly known
  as diagrammatic modeling language (\eg, visualizing class models),
  it is complemented by a textual language, called Object Constraint
  Language (OCL). OCL is a textual annotation language, based on a
  three-valued logic, that turns UML into a formal language.
  Unfortunately the semantics of this specification language, captured
  in the ``Annex A'' of the OCL standard, leads to different
  interpretations of corner cases.  Many of these corner cases had
  been subject to formal analysis since more than ten years.

  The situation complicated when with version 2.3 the OCL was aligned
  with the latest version of UML: this led to the extension of the
  three-valued logic by a second exception element, called
  \inlineocl{null}.  While the first exception element
  \inlineocl{invalid} has a strict semantics, \inlineocl{null} has a
  non strict semantic interpretation. These semantic difficulties lead
  to remarkable confusion for implementors of OCL compilers and
  interpreters.

  In this paper, we provide a formalization of the core of OCL in
  HOL\@. It provides denotational definitions, a logical calculus and
  operational rules that allow for the execution of OCL expressions by
  a mixture of term rewriting and code compilation.  Our formalization
  reveals several inconsistencies and contradictions in the current
  version of the OCL standard.  They reflect a challenge to define and
  implement OCL tools in a uniform manner.  Overall, this document is
  intended to provide the basis for a machine-checked text ``Annex A''
  of the OCL standard targeting at tool implementors.

\end{abstract}

\tableofcontents
\include{introduction}
\include{formalization}
\include{conclusion}
\bibliographystyle{abbrvnat}
\bibliography{root} 

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%  LocalWords:  implementors denotational OCL UML
